---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Code for analyses of EXPERIMENT 1 (reproductive tradeoffs) reported in Garcia and Orians (2022). "Reproductive tradeoffs in a perennial crop: Exploring the mechanisms of coffee alternate bearing in relation to farm management." Agriculture, Ecosystems, and Environment.


```{r}
library(ggeffects)
library(Hmisc)
library(glm2)
library(glmmTMB)
library(dplyr)
library(tidyverse)
library(lme4)
library(boot)
library(bbmle)
library(zoo)
library(car)
library(lmtest)
library(mgcv)
library(pscl)
library(multcomp)
library(emmeans)
library(ggsignif)
library(ggpubr)
library(ggplot2)
library(MuMIn)

```



# read in clean 2019 data and mutate IDs to factors
```{r}

ndat <- read.csv( "cleandat_n1i_n1h.csv", header = T)
head(ndat)

ndat <- ndat %>%
  mutate(branch_id = as.factor(branch_id)) %>% 
  mutate(plant_id = as.factor(plant_id)) %>% 
  mutate(per_thin = as.factor(per_thin)) %>% 
  mutate(n1i_frt = as.integer(n1i_frt))

```



# read in 2018 data and clean up (code to make sure numbers are read as numbers)

```{r}
edat2h = read.csv("manip_e1i_e2f_e2i_e2h.csv", header=T)

edath_fix<- edat2h %>% 
  mutate(e2h_tot_len = as.character(e2h_tot_len)) %>% 
  mutate(e2h_new_len= as.character(e2h_new_len)) %>% 
  mutate(e2h_diam = as.character(e2h_diam)) %>% 
  mutate(e2h_tot_nod = as.character(e2h_tot_nod)) %>% 
  mutate(e2h_new_nod= as.character(e2h_new_nod)) %>% 
  mutate(e2h_frtd_nod = as.character(e2h_frtd_nod)) %>% 
  mutate(e2h_frt = as.character(e2h_frt)) %>% 
  mutate(e2h_tot_lvs= as.character(e2h_tot_lvs)) %>% 
  mutate(e2h_sick_lvs = as.character(e2h_sick_lvs)) 

edath_full <-   edath_fix %>% 
  mutate(e2h_tot_len = as.numeric(e2h_tot_len)) %>% 
  mutate(e2h_new_len= as.numeric(e2h_new_len)) %>% 
  mutate(e2h_diam = as.numeric(e2h_diam)) %>% 
  mutate(e2h_tot_nod = as.numeric(e2h_tot_nod)) %>% 
  mutate(e2h_new_nod= as.numeric(e2h_new_nod)) %>% 
  mutate(e2h_frtd_nod = as.numeric(e2h_frtd_nod)) %>% 
  mutate(e2h_frt = as.numeric(e2h_frt)) %>% 
  mutate(e2h_tot_lvs= as.numeric(e2h_tot_lvs)) %>% 
  mutate(e2h_sick_lvs = as.numeric(e2h_sick_lvs)) 

head(edath_full)

edath_full = edath_full %>%  
  mutate(branch_id = as.factor(branch_id)) %>% 
  mutate(plant_id = as.factor(plant_id)) %>% 
  mutate(per_thin = as.factor(per_thin))

head(edath_full)

```


# Models included in the body of the manuscript.

# 1 Initiated fruited nodes - e2i (neg binom)

## 1.1 model 

```{r}
head(edath_full)

range(edat$e2i_frt_nod, na.rm = T)
range(ndat$n1i_frtd_nod, na.rm = T)

# compare poisson and negative binomial fits 

#m_fruitnod = glmer(e2i_frt_nod~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), family=poisson(link = "log"), data=edat)

m_fruitnod_nb = glmer.nb(e2i_frt_nod~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), data=edath_full)

#ICtab(m_fruitnod, m_fruitnod_nb)

summary(m_fruitnod_nb)
Anova(m_fruitnod_nb)
as.data.frame(Anova(m_fruitnod_nb))

# checking if fixed effect of farmer changes things
m_fruitnod_nb_f = glmer.nb(e2i_frt_nod~per_thin*fert_trt*shd_trt+ farmer+(1|plant_id), data=edath_full)

# post hocs 

## thinning 
frtnod_thin_post = emmeans(m_fruitnod_nb,  pairwise ~ per_thin, type="response", adjust="none")
eifrtnod_thin_ci <- as.data.frame(frtnod_thin_post$emmeans)

## fertilizer
frtnod_fert_post = emmeans(m_fruitnod_nb,  pairwise ~ fert_trt, type="response", adjust="none")
eifrtnod_fert_ci <- as.data.frame(frtnod_fert_post$emmeans)

## shade
frtnod_shd_post = emmeans(m_fruitnod_nb,  pairwise ~ shd_trt, type="response", adjust="none")


## fert-shd for plot 2A

frtnod_fertshd = emmeans(m_fruitnod_nb,  pairwise ~ fert_trt*shd_trt, type="response", adjust="none")
eifrtnod_fertshd_ci <- as.data.frame(frtnod_fertshd$emmeans)

```

## 1.2 thinning plot - e2i fnod (PLOT 2A)

```{r}

ggplot(eifrtnod_thin_ci, aes(per_thin, response, fill=per_thin))+
    geom_errorbar(data = eifrtnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "Old Branch:
  Initiated Fruited Nodes", limits = c(2,8), breaks = c(2,4,6,8))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")

```

## 1.3 fert-shd plot (PLOT 3A)

```{r}
ggplot(eifrtnod_fertshd_ci, aes(x=shd_trt, y=response, group=fert_trt, color=fert_trt, fill=fert_trt))+
    geom_errorbar(data = eifrtnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "Old Branch:
  Initiated Fruited Nodes", limits = c(2,8))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))

```

# 2.  Initiated fruits per fruited node - e2i (neg binom) 

# 2.1 model - e2i ffnod

```{r}
# model comparison - poisson vs. negative binomial

#m_initfrt <- glmer(e2i_frts~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(e2i_frt_nod), family=poisson(link = "log"), data=edat[edat$e2i_frt_nod >0,])
m_initfrt_nb <- glmer.nb(e2i_frts~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(e2i_frt_nod),data=edat[edat$e2i_frt_nod >0,])
#ICtab(m_initfrt, m_initfrt_nb)

summary(m_initfrt_nb)
Anova(m_initfrt_nb)
as.data.frame(Anova(m_initfrt_nb))

# post hocs

# thinning
ffnod_thin_post = emmeans(m_initfrt_nb, pairwise ~ per_thin, adjust="none", type="response")
effnod_thin_ci = as.data.frame(ffnod_thin_post$emmeans)
# shade
ffnod_shd_post = emmeans(m_initfrt_nb, pairwise ~ shd_trt, adjust="none", type="response")
effnod_shd_ci= as.data.frame(ffnod_shd_post$emmeans)
# fertilizer 
ffnod_fert_post = emmeans(m_initfrt_nb, pairwise ~ fert_trt, adjust="none", type="response")

# fert shd plot 
ffnod_fertshd_post = emmeans(m_initfrt_nb, pairwise ~ fert_trt*shd_trt, adjust="none", type="response")
effnod_fertshd_ci= as.data.frame(ffnod_fertshd_post$emmeans)
```

## 2.2 thin plot (PLOT 2B)

```{r}

ggplot(effnod_thin_ci, aes(per_thin, response, fill=per_thin))+
    geom_errorbar(data = effnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "Old Branch:
  Initiated Fruits per Node", limits = c(2,8), breaks = c(4,6,8,10,12))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")

```


## 2.3 fert shd plot - e2i ffnod (PLOT 3B)
```{r}

ggplot(effnod_fertshd_ci, aes(shd_trt, response,group=fert_trt, color=fert_trt, fill=fert_trt ))+
  geom_errorbar(data = effnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "Old Branch:
  Initiated Fruits per Node", limits = c(2,8))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))

```


# 3. Initiated fruited nodes, new cohort (n1i) - poisson

## 3.1 model 

```{r}
# model comparison - poisson vs. negative binomial

m_fruitnod_n1 = glmer(n1i_frtd_nod~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), family=poisson(link = "log"), data=ndat)

#m_fruitnod_n1_nb = glmer.nb(n1i_frtd_nod~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), data=ndat)

#ICtab(m_fruitnod_n1, m_fruitnod_n1_nb)

summary(m_fruitnod_n1)
Anova(m_fruitnod_n1)
as.data.frame(Anova(m_fruitnod_n1))

# post hocs

## thinning 
n1_frtnod_thin_post = emmeans(m_fruitnod_n1,  pairwise ~ per_thin, type="response", adjust="none")
nifnod_thin_ci = as.data.frame(n1_frtnod_thin_post$emmeans)

## fertilizer
n1_frtnod_fert_post = emmeans(m_fruitnod_n1,  pairwise ~ fert_trt, type="response", adjust="none")
nifnod_fert_ci = as.data.frame(n1_frtnod_fert_post$emmeans)

## 3 way
n1_frtnod_fert_post = emmeans(m_fruitnod_n1,  pairwise ~ per_thin*fert_trt*shd_trt, type="response", adjust="none")
nifnod_3_ci = as.data.frame(n1_frtnod_fert_post$emmeans)

# shade 
n1_frtnod_shd_post = emmeans(m_fruitnod_n1,  pairwise ~ shd_trt, type="response", adjust="none")

#fert-shade  
n1_frtshd_fert_post = emmeans(m_fruitnod_n1,  pairwise ~ fert_trt*shd_trt, type="response", adjust="none")
nifnod_fertshd_ci = as.data.frame(n1_frtshd_fert_post$emmeans)
```


## 3.2 thin plot - n1i fnod (PLOT 2E)
```{r}
ggplot(nifnod_thin_ci, aes(per_thin, rate, fill=per_thin))+
    geom_errorbar(data = nifnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "New Branch:
  Initiated Fruited Nodes", limits = c(2,12), breaks = c(2,4,6,8,10,12))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")

```


## 3.3 nifnod fert shade plot (PLOT 3E)
```{r}

ggplot(nifnod_fertshd_ci, aes(shd_trt, rate,group=fert_trt, color=fert_trt, fill=fert_trt ))+
  geom_errorbar(data = nifnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "New Branch:
  Initiated Fruited Nodes", limits = c(2,12), breaks = c(2,4,6,8,10,12))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))

```

## 3.4 Appendix 3:  3-way plot - n1i fnod

```{r}
nifnod_3_ci$plot_thin <- paste(nifnod_3_ci$per_thin, nifnod_3_ci$fert_trt, nifnod_3_ci$shd_trt, sep=" ")

ggplot(nifnod_3_ci, aes(x=plot_thin, y=rate))+
  geom_point(aes(fill=plot_thin), colour="black", size=4, pch=21)+
  geom_errorbar(data=nifnod_3_ci, mapping=aes(ymin=asymp.LCL, ymax=asymp.UCL), width=0.5, color="black")+
  scale_x_discrete(name="",limits= c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), labels = c("33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%"))+
  scale_fill_manual(limits = c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), values = c("#662506","#993404", "#cc4c02","#004529","#006837", "#238443", "#ec7014" , "#fe9929", "#fec44f", "#41ab5d", "#78c679","#addd8e"))+
  theme_bw()+
  theme(text = element_text(size = 16), legend.position = "none")+
  scale_y_continuous(name = "New Branch: Initiated Fruited Nodes",  limits=c(4, 10), breaks=c(6, 8, 10))

```



# 4. Initiated fruits per fruited node - new cohort (n1i) - neg binom

## 4.1 model 

```{r}

#m_initfrt_n1 <- glmer(n1i_frt~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(n1i_frtd_nod), family=poisson(link = "log"), data=ndat[ndat$n1i_frtd_nod >0,])

m_initfrt_n1_nb <- glmer.nb(n1i_frt~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(n1i_frtd_nod), data=ndat[ndat$n1i_frtd_nod >0,])

#ICtab(m_initfrt_n1, m_initfrt_n1_nb)

summary(m_initfrt_n1_nb)
Anova(m_initfrt_n1_nb)
as.data.frame(Anova(m_initfrt_n1_nb))

# post hocs

# thinning 
n1_frt_thin_post = emmeans(m_initfrt_n1_nb,  pairwise ~ per_thin, type="response", adjust="none")
niffnod_thin_ci = as.data.frame(n1_frt_thin_post$emmeans)

# fert 
n1_frt_fert_post = emmeans(m_initfrt_n1_nb,  pairwise ~ fert_trt, type="response", adjust="none")
niffnod_fert_ci = as.data.frame(n1_frt_fert_post$emmeans)

# shade 
n1_frt_shd_post = emmeans(m_initfrt_n1_nb,  pairwise ~ shd_trt, type="response", adjust="none")
niffnod_shd_ci= as.data.frame(n1_frt_shd_post$emmeans)

# fert shd plot 
n1ffnod_fertshd_post = emmeans(m_initfrt_n1_nb,  pairwise ~ fert_trt*shd_trt, type="response", adjust="none")
niffnod_fertshd_ci= as.data.frame(n1ffnod_fertshd_post$emmeans)


```

## 4.2 thin plot - n1i ffnod (PLOT 2F)
```{r}
ggplot(niffnod_thin_ci, aes(per_thin, response, fill=per_thin))+
  geom_errorbar(data = niffnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "New Branch:
  Initiated Fruits per Node", limits = c(4,12), breaks = c(2,4,6,8,10, 12))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")


```


## 4.3 fert shd n1i ffnod (PLOT 3F)

```{r}
ggplot(niffnod_fertshd_ci, aes(shd_trt, response, group=fert_trt, color=fert_trt, fill=fert_trt ))+
  geom_errorbar(data = niffnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "New Branch:
  Initiated Fruits per Node", limits = c(2,12), breaks = c(2,4,6,8,10,12))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))
```



# 5. Mature fruited nodes - old cohort (e2h)- neg binom

## 5.1 model

```{r}


#m_harfrtdnod <- glmer(e2h_frtd_nod~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), family=poisson(link = "log"), data=edath_full)
m_harfrtdnod_nb <- glmer.nb(e2h_frtd_nod~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), data=edath_full)
#ICtab(m_harfrtdnod, m_harfrtdnod_nb)

summary(m_harfrtdnod_nb)
Anova(m_harfrtdnod_nb)
as.data.frame(Anova(m_harfrtdnod_nb))

# thinning 
harfrtdnod_thin_post = emmeans(m_harfrtdnod_nb,  pairwise ~ per_thin, type="response", adjust="none")
e2hfnod_thin_ci= as.data.frame(harfrtdnod_thin_post$emmeans)

#fert 
harfrtdnod_fert_post = emmeans(m_harfrtdnod_nb,  pairwise ~ fert_trt, type="response", adjust="none")
e2hfnod_fert_ci = as.data.frame(harfrtdnod_fert_post$emmeans)

# shade 
harfrtdnod_shd_post = emmeans(m_harfrtdnod_nb,  pairwise ~ shd_trt, type="response", adjust="none")
e2hfnod_shd_ci= as.data.frame(harfrtdnod_shd_post$emmeans)

# fert shd  
harfrtdnod_fertshd_post = emmeans(m_harfrtdnod_nb,  pairwise ~ fert_trt*shd_trt, type="response", adjust="none")
e2hfnod_fertshd_ci= as.data.frame(harfrtdnod_fertshd_post$emmeans)
```


## 5.2 thin plot - e2h fnod (PLOT 2C)

```{r}
ggplot(e2hfnod_thin_ci, aes(per_thin, response, fill=per_thin))+
  geom_errorbar(data = e2hfnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")+
  scale_y_continuous( name = "Old Branch:
  Mature Fruited Nodes", limits = c(2,8), breaks = c(2,4,6,8,10,12))


```


## 5.3 fert shd e2h fnod (PLOT 3C)

```{r}
ggplot(e2hfnod_fertshd_ci, aes(shd_trt, response, group=fert_trt, color=fert_trt, fill=fert_trt ))+
  geom_errorbar(data = e2hfnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "Old Branch:
  Mature Fruited Nodes", limits = c(2,8), breaks = c(2,4,6,8,10,12))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))

```

# 6. Mature fruits per fruited node - old cohort (e2h)- neg binom

## 6.1 model 
```{r}

#m_harffnod <- glmer(e2h_frt~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(e2h_frtd_nod), family=poisson(link = "log"), data=edath_full[edath_full$e2h_frtd_nod >0,])

m_harffnod_nb <- glmer.nb(e2h_frt~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(e2h_frtd_nod), data=edath_full[edath_full$e2h_frtd_nod >0,])

#ICtab(m_harffnod, m_harffnod_nb)

summary(m_harffnod_nb)
Anova(m_harffnod_nb)
as.data.frame(Anova(m_harffnod_nb))

# thin  
harffnod_thin_post = emmeans(m_harffnod_nb,  pairwise ~ per_thin, type="response", adjust="none")
ehffnod_thin_ci = as.data.frame(harffnod_thin_post$emmeans)

# shade 
harffnod_shd_post = emmeans(m_harffnod_nb,  pairwise ~ shd_trt, type="response", adjust="none")
ehffnod_shd_ci = as.data.frame(harffnod_shd_post$emmeans)

# fert 
harffnod_fert_post = emmeans(m_harffnod_nb,  pairwise ~ fert_trt, type="response", adjust="none")

# fert shd plot 
harffnod_fertshd_post = emmeans(m_harffnod_nb,  pairwise ~ fert_trt*shd_trt, type="response", adjust="none")
ehffnod_fertshd_ci = as.data.frame(harffnod_fertshd_post$emmeans)
```

## 6.2 thin plot - - e2h ffnod (PLOT 2D)

```{r}
ggplot(ehffnod_thin_ci, aes(per_thin, response, fill=per_thin))+
  geom_errorbar(data = ehffnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")+
  scale_y_continuous( name = "Old Branch:
  Mature Fruits per Node", limits = c(2,8), breaks = c(2,4,6,8))

```


## 6.3 e2h ffond fert shd plot (PLOT 3D)
```{r}
ggplot(ehffnod_fertshd_ci, aes(shd_trt, response, group=fert_trt, color=fert_trt, fill=fert_trt ))+
  geom_errorbar(data = ehffnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "Old Branch:
  Mature Fruits per Node", limits = c(2,8), breaks = c(2,4,6,8,10,12))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))

```


# 7. Mature fruited nodes - new branch cohort (n1h)- poisson 

## 7.1 model
```{r}
m_harfrtdnod_n1 <- glmer(n1h_frtd_nod~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), family=poisson(link = "log"), data=ndat)

#m_harfrtdnod_n1_nb <- glmer.nb(n1h_frtd_nod~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id),  data=ndat)

#ICtab(m_harfrtdnod_n1, m_harfrtdnod_n1_nb)

Anova(m_harfrtdnod_n1)
as.data.frame(Anova(m_harfrtdnod_n1))

#thin - non sig! - supplement
n1_harfrtdnod_thin_post = emmeans(m_harfrtdnod_n1, pairwise ~ per_thin, type ="response", adjust="none")
n1ihfnod_thin_ci = as.data.frame(n1_harfrtdnod_thin_post$emmeans)

# fert 
n1_harfrtdnod_fert_post = emmeans(m_harfrtdnod_n1, pairwise ~ fert_trt, type ="response", adjust="none")
n1hfnod_fert_ci = as.data.frame(n1_harfrtdnod_fert_post$emmeans)


# 3 way - supplement
n1_harfrtdnod_three_post = emmeans(m_harfrtdnod_n1, pairwise ~ per_thin*fert_trt*shd_trt, type ="response", adjust="none")
n1hfnod_3_ci = as.data.frame(n1_harfrtdnod_three_post$emmeans)

# shade 
n1_harfrtdnod_shd_post = emmeans(m_harfrtdnod_n1, pairwise ~ shd_trt, type ="response", adjust="none")

# fert shd 
n1_harfrtdnod_fertshd_post = emmeans(m_harfrtdnod_n1, pairwise ~ fert_trt*shd_trt, type ="response", adjust="none")
n1hfnod_fertshd_ci = as.data.frame(n1_harfrtdnod_fertshd_post$emmeans)

```

## 7.2 thin plot - n1h fnod (not sig.)

```{r}
ggplot(n1ihfnod_thin_ci, aes(per_thin, rate, fill=per_thin))+
  geom_errorbar(data = n1ihfnod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")+
  scale_y_continuous( name = "New Branch:
  Mature Fruited Nodes", limits = c(2,9), breaks = c(2,4,6,8,10,12))

```

## 7.3 n1h fnod fert shd plot (PLOT 3G)

```{r}
ggplot(n1hfnod_fertshd_ci, aes(shd_trt, rate, group=fert_trt, color=fert_trt, fill=fert_trt ))+
  geom_errorbar(data = n1hfnod_fertshd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black", position = position_dodge(width = 0.5))+
  geom_point(size=4,  color="black", pch=21, position = position_dodge(width = 0.5))+
  geom_line( position = position_dodge(width = 0.5))+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "New Branch:
  Mature Fruited Nodes", limits = c(2,8.5), breaks = c(2,4,6,8,10,12))+
  scale_color_manual(name="Fertilizer",values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  scale_fill_manual(name="Fertilizer", values= c("coral", "cornflowerblue"), labels=c("high", "low"))+
  theme_bw()+
  theme(text = element_text(size = 20))

```


## 7.4 Appendix 3. plot 3 way - n1h fnod

```{r}
n1hfnod_3_ci$plot_thin <- paste(n1hfnod_3_ci$per_thin, n1hfnod_3_ci$fert_trt, n1hfnod_3_ci$shd_trt, sep=" ")


ggplot(n1hfnod_3_ci, aes(x=plot_thin, y=rate))+
  geom_point(aes(fill=plot_thin), colour="black", size=4, pch=21)+
  geom_errorbar(data=n1hfnod_3_ci, mapping=aes(ymin=asymp.LCL, ymax=asymp.UCL), width=0.5, color="black")+
  scale_x_discrete(name="",limits= c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), labels = c("33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%"))+
  scale_fill_manual(limits = c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), values = c("#662506","#993404", "#cc4c02","#004529","#006837", "#238443", "#ec7014" , "#fe9929", "#fec44f", "#41ab5d", "#78c679","#addd8e"))+
  theme_bw()+
  theme(text = element_text(size = 16), legend.position = "none")+
  scale_y_continuous(name = "New cohort: Mature Fruited Nodes",  limits=c(4, 10), breaks=c(6, 8, 10))

```


#8. Mature fruits per fruited node - new branch cohort (n1h)- neg binom

## 8.1 model 

```{r}
#m_harffnod_n1 <- glmer(n1h_frt~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(n1h_frtd_nod), family=poisson(link = "log"), data=ndat[ndat$n1h_frtd_nod >0,])

m_harffnod_n1_nb <- glmer.nb(n1h_frt~per_thin*fert_trt*shd_trt +(1|farmer)+(1|plant_id), offset = log(n1h_frtd_nod), data=ndat[ndat$n1h_frtd_nod >0,])

#ICtab(m_harffnod_n1, m_harffnod_n1_nb)

summary(m_harffnod_n1_nb)
Anova(m_harffnod_n1_nb)

# thin 
n1_harffnod_thin_post = emmeans(m_harffnod_n1_nb, pairwise ~ per_thin, type ="response", adjust="none")
n1ihffnod_thin_ci = as.data.frame(n1_harffnod_thin_post$emmeans)

# fert
n1_harffnod_fert_post = emmeans(m_harffnod_n1_nb, pairwise ~ fert_trt, type ="response", adjust="none")

# shade 
n1_harffnod_shd_post = emmeans(m_harffnod_n1_nb, pairwise ~ shd_trt, type ="response", adjust="none")


```


# 9. Cumulative average fruits per branch in old cohort across years. 

## data prep
```{r}

# making a treatment dataframe to merge
edat_trt <- edath_full %>% 
  dplyr::group_by(plant_id) %>% 
  dplyr::select(farmer,fert_trt, shd_trt, per_thin)

edat_trt_unique = distinct(edat_trt, plant_id, .keep_all = TRUE)

# average fruits  in y (y cohort)
edat1h_avgfrts <- edath_full %>%
  group_by(plant_id) %>% 
  dplyr::summarize(e1h_avg_frts= mean(e1h_tot_fruit, na.rm = TRUE)) 

# average fruits in y+1 (y cohort)
edat2h_avgfrts <- edath_full %>%
  group_by(plant_id) %>% 
  dplyr::summarize(e2h_avg_frts= mean(e2h_frt, na.rm = TRUE)) 

e12_sumfrts <-  merge(edat1h_avgfrts, edat2h_avgfrts, by = "plant_id")
e12_sum_trt <- merge(e12_sumfrts, edat_trt_unique,  by ="plant_id")

# looking at y and y+1 sums within the y cohort 

e12_sum_trt$sum_12hfrts <- e12_sum_trt$e1h_avg_frts + e12_sum_trt$e2h_avg_frts



```


## model

```{r}
# mh_12sumffnod0 <- glmer(as.integer(sum_12hfrts) ~ per_thin * fert_trt * shd_trt +(1|farmer), family = poisson(), data = e12_sums)

mh_12sumfrts <- glmer.nb(as.integer(sum_12hfrts) ~ per_thin * fert_trt * shd_trt +(1|farmer),  data = e12_sum_trt)


#ICtab(mh_12sumffnod0, mh_12sumffnodnb)

summary(mh_12sumfrts)
Anova(mh_12sumfrts)

emmeans(mh_12sumfrts, pairwise ~ per_thin, type="response", adjust="none")
emmeans(mh_12sumfrts, pairwise ~ fert_trt*shd_trt, type = "response", adjust="none")
```

#########################################################


# Assessment of pre-existing differences across treatments 


## 0A. total nodes 

pre_totnod = pre-existing total nodes per branch
```{r}
pre_totnod <- glmer(e1i_tot_nod~per_thin+fert_trt+shd_trt +(1|farmer)+(1|plant_id), data=edath_full, family=poisson(link = "log"))
Anova(pre_totnod)

emmeans(pre_totnod, pairwise ~ shd_trt, type="response", adjust = "none")

```


## 0B. Initiated  fruits

pre_prefrt = pre-existing fruits per branch. 
e1i_pot_nod = 2018 branch, 1st year of production, i = fruit initation; pot_nod = nodes with fruiting potential (not the newly extended nodes)

```{r}
# pre thin
pre_prefrt<- glmer(e1i_pre_frt~per_thin+fert_trt+shd_trt +(1|farmer)+(1|plant_id), offset = log(e1i_pot_nod), data=edath_full[edath_full$e1i_pot_nod>0,], family=poisson(link = "log"))
  
summary(pre_prefrt)
Anova(pre_prefrt)

```

## 0C. nodes with fruiting potential

```{r}
pre_potnod <-glmer(e1i_pot_nod ~ per_thin + fert_trt + shd_trt + (1|farmer) + (1|plant_id), family = poisson (link = "log"), data= edath_full)
Anova(pre_potnod)

emmeans(pre_potnod, pairwise ~ shd_trt, type="response", adjust="none")

```


## 0D. leaves 

```{r}

pre_leaves <-glmer(e1i_tot_lvs ~ per_thin + fert_trt + shd_trt +(1|farmer) + (1|plant_id), family = poisson(link = "log"), data= edath_full)
Anova(pre_leaves)
emmeans(pre_leaves, pairwise ~ shd_trt, type="response", adjust="none")

```

## 0E. length 

```{r}
pre_length <- lmer(e1i_len~per_thin+fert_trt+shd_trt + (1|farmer)+(1|plant_id), data=edath_full)
Anova(pre_length)
emmeans(pre_length, pairwise ~ shd_trt, adjust="none")
```


# Appendix 1 

analyze growth, fruited nodes, and fruits per fruited node at maturity in y (2018) and flowering nodes and buds per node at flowering in 2019. 

##  branch growth measured at fruit maturity in y (e1har)
```{r}
# make growth variable
edath_full$e1h_growth <- edath_full$e1h_len - edath_full$e1i_len
hist(edath_full$e1h_growth)

# growth model 
ehar_growth <- lmer(e1h_growth~per_thin*fert_trt*shd_trt + (1|farmer)+(1|plant_id), data=edath_full)
Anova(ehar_growth)
summary(ehar_growth)

# post hoc and confidence intervals for plots
ehar_growth_post <- emmeans(ehar_growth, pairwise ~ per_thin, adjust="none")
ehgrowth_thin_ci <- as.data.frame(ehar_growth_post$emmeans)


### percent thin plot
ggplot(ehgrowth_thin_ci, aes(per_thin, emmean, fill=per_thin))+
    geom_errorbar(data = ehgrowth_thin_ci, mapping = aes(ymin=lower.CL, ymax=upper.CL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "Branch growth in cm (y)", limits = c(2,16))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")




### 3-way plot
ehar_growth_3way_post <- emmeans(ehar_growth, pairwise ~ per_thin*fert_trt*shd_trt, adjust="none")
ehgrowth_3way_ci <- as.data.frame(ehar_growth_3way_post$emmeans)

ehgrowth_3way_ci$plot_thin <- paste(ehgrowth_3way_ci$per_thin, ehgrowth_3way_ci$fert_trt, ehgrowth_3way_ci$shd_trt, sep=" ")

ggplot(ehgrowth_3way_ci, aes(x=plot_thin, y=emmean))+
  geom_point(aes(fill=plot_thin), colour="black", size=4, pch=21)+
  geom_errorbar(data=ehgrowth_3way_ci, mapping=aes(ymin=lower.CL, ymax=upper.CL), width=0.5, color="black") +
  scale_fill_manual(limits = c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), values = c("#662506","#993404", "#cc4c02","#004529","#006837", "#238443", "#ec7014" , "#fe9929", "#fec44f", "#41ab5d", "#78c679","#addd8e"))+
  scale_x_discrete(name="",limits= c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), labels = c("33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%"))+
  theme_bw()+
  theme(text = element_text(size = 16), legend.position = "none")  +
scale_y_continuous(name = "Branch growth in cm (y)",  limits=c(2, 16))

```


## floral nodes in y+1 (e2f)


```{r}
head(edath_full)
#m_flornod = glmer(e2f_flor_nod~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), family=poisson(link = "log"), data=edath_full)
edath_full$e2i_frt_nod
m_flornod_nb = glmer.nb(e2f_flor_nod~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), data=edath_full[edath_full$e2f_flor_nod >0, ])
ICtab(m_flornod, m_flornod_nb)

summary(m_flornod_nb)
Anova(m_flornod_nb)

## thinning post hoc
flornod_thin_post = emmeans(m_flornod_nb,  pairwise ~ per_thin, type="response", adjust="none")
e1flornod_thin_ci <- as.data.frame(flornod_thin_post$emmeans)

### thin plot
ggplot(e1flornod_thin_ci, aes(per_thin, response, fill=per_thin))+
    geom_errorbar(data = e1flornod_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "Floral nodes (y+1)", limits = c(2,11))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")



## fertilizer post hoc 
flornod_fert_post = emmeans(m_flornod_nb,  pairwise ~ fert_trt, type="response", adjust="none")
e1flornod_fert_ci <- as.data.frame(flornod_fert_post$emmeans)


## three way post hoc
flornod_3way_post = emmeans(m_flornod_nb,  pairwise ~ per_thin*fert_trt*shd_trt, type="response", adjust="none")
e2flornod_3way_ci <- as.data.frame(flornod_3way_post$emmeans)


e2flornod_3way_ci$plot_thin <- paste(e2flornod_3way_ci$per_thin, e2flornod_3way_ci$fert_trt, e2flornod_3way_ci$shd_trt, sep=" ")

ggplot(e2flornod_3way_ci, aes(x=plot_thin, y=response))+
  geom_point(aes(fill=plot_thin), colour="black", size=4, pch=21)+
  geom_errorbar(data=e2flornod_3way_ci, mapping=aes(ymin=asymp.LCL, ymax=asymp.UCL), width=0.5, color="black") +
  scale_fill_manual(limits = c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), values = c("#662506","#993404", "#cc4c02","#004529","#006837", "#238443", "#ec7014" , "#fe9929", "#fec44f", "#41ab5d", "#78c679","#addd8e"))+
  scale_x_discrete(name="",limits= c("66 h sun","33 h sun", "0 h sun",  "66 h shd", "33 h shd", "0 h shd", "66 l sun", "33 l sun", "0 l sun",  "66 l shd", "33 l shd", "0 l shd"), labels = c("33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%","33%", "66%", "100%"))+
  theme_bw()+
  theme(text = element_text(size = 16), legend.position = "none")  +
scale_y_continuous(name = "Floral nodes (y+1)",  limits=c(2, 11))


```


## e2f floral buds per node 


```{r}
head(edath_full)
#m_florbud = glmer(e2f_buds~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id), offset=log(e2f_flor_nod), family=poisson(link = "log"), data=edath_full[edath_full$e2f_buds >0, ])

m_florbud_nb = glmer.nb(e2f_buds~per_thin*fert_trt*shd_trt+(1|farmer)+(1|plant_id),offset=log(e2f_flor_nod), data=edath_full[edath_full$e2f_buds >0, ])
#ICtab(m_florbud, m_florbud_nb)

Anova(m_florbud_nb)


## thinning post hoc
florbud_thin_post = emmeans(m_florbud_nb,  pairwise ~ per_thin, type="response", adjust="none")
e2florbud_thin_ci <- as.data.frame(florbud_thin_post$emmeans)

### thin plot
ggplot(e2florbud_thin_ci, aes(per_thin, response, fill=per_thin))+
    geom_errorbar(data = e2florbud_thin_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "33", "0"), labels = c("33%", "66%", "100%"))+
  scale_y_continuous(name = "Floral buds per node (y+1)", limits = c(2,11))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Greens",direction = -1)+
  theme(legend.position = "none")


## shade  post hoc
florbud_shd_post = emmeans(m_florbud_nb,  pairwise ~ shd_trt, type="response", adjust="none")
e2florbud_shd_ci <- as.data.frame(florbud_shd_post$emmeans)

### shade plot
ggplot(e2florbud_shd_ci, aes(shd_trt, response, fill=shd_trt))+
    geom_errorbar(data = e2florbud_shd_ci, mapping = aes(ymin=asymp.LCL, ymax=asymp.UCL), width = 0.5, color="black")+
  geom_point(size=5, color="black", pch=21)+
  scale_x_discrete(name="Light Treatment", limits=c("sun", "shd"), labels=c("sun-grown", "shaded"))+
  scale_y_continuous(name = "Floral buds per node (y+1)", limits = c(2,11))+
  theme_bw()+
  theme(text = element_text(size = 20))+
  scale_fill_brewer(palette = "Reds",direction = 1)+
  theme(legend.position = "none")


```


#########################################################


