---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Code for analyses of EXPERIMENT 2 (stem nitrogen) reported in Garcia and Orians (2022). "Reproductive tradeoffs in a perennial crop: Exploring the mechanisms of coffee alternate bearing in relation to farm management." Agriculture, Ecosystems, and Environment.


# install / load packages 


```{r}
library(dplyr)
library(tidyverse)
library(lme4)
library(boot)
library(emmeans)
library(car)
```


# MATURITY 2019 (y)

## load data
I referred to maturity as fruit harvest - so this section has "h"'s and "har"'s to distinguish it from later sampling points. 
```{r}


h19res <- read.csv("Harvest2019_Results - clean.csv", header = T)

head(h19res) 

```

# adjust the data
make sample id and thinning a discrete factor rather than a number. 
```{r}

har19res<- h19res %>%
  mutate(per_thin = as.factor(per_thin)) %>% 
  mutate(sample_id = as.factor(sample_id))


```

# subset stem data
```{r}
har19stem <- har19res[har19res$tissue =="stem",]
head(har19stem)

```

###  model 
```{r}
h19stem_N <- glm(per_n ~ per_thin * shd_trt * fert_trt, data = har19stem)
summary(h19stem_N)
Anova(h19stem_N)

as.data.frame(Anova(h19stem_N))
har19_stemN_thin <- emmeans(h19stem_N, pairwise ~ per_thin, adjust="none")

# post hocs
emmeans(h19stem_N, pairwise ~ per_thin*shd_trt,adjust="none")
emmeans(h19stem_N, pairwise ~shd_trt,adjust="none")
emmeans(h19stem_N, pairwise ~fert_trt,adjust="none")

h19stem_thin_ci = as.data.frame(har19_stemN_thin$emmeans)
```

###  plot 

```{r}
ggplot(h19stem_thin_ci, aes(per_thin, emmean, fill=per_thin))+
  geom_errorbar(data = h19stem_thin_ci, mapping = aes(ymin=lower.CL, ymax=upper.CL), width = 0.5, color="black")+
    geom_point(size=4, pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "0"), labels = c("33%","100%"))+
  scale_y_continuous(name = "Stem % N at Fruit Maturity (y)", limits = c(1,2))+
  theme_bw()+
  scale_fill_manual(values = c( "#2ca25f", "#e5f5f9"))+
  theme(text = element_text(size = 16))+
  theme(legend.position = "none")

```


# FLOWERING 2020 (y+1)

```{r}

fl20res <- read.csv("Resource Analysis_Floral_2020.csv", header=T)

head(fl20res)
```

```{r}
fl20res<- fl20res %>%  
  mutate(per_thin = as.factor(per_thin)) %>% 
  mutate(sample_id = as.factor(sample_id))
head(fl20res)
```

```{r}
f20stem <- fl20res[fl20res$tissue=="stem",]
head(f20stem)
```


### model 

```{r}
f20stem_N <- lm(per_n ~ per_thin * shd_trt * fert_trt, data = f20stem)
summary(f20stem_N)
Anova(f20stem_N)
as.data.frame(Anova(f20stem_N))

fl20_stN_thin <- emmeans(f20stem_N, pairwise ~ per_thin, adjust="none")
f20stemN_thin_ci = as.data.frame(fl20_stN_thin$emmeans)
```




### plot

# FRUIT INITIATION 2020 (y+1)

```{r}

i20res <- read.csv("Resource Analysis_Init_2020.csv", header = T)

#inspect the first six rows, make sure it looks good. 
head(i20res) 

```

```{r}

init20res<- i20res %>%
  mutate(per_thin = as.factor(per_thin)) %>% 
  mutate(sample_id = as.factor(sample_id))


```

```{r}
init20stem <- init20res[init20res$tissue =="stem",]
 head(init20stem)

```

### model 
```{r}
i20stem_N <- lm(per_n ~ per_thin * shd_trt * fert_trt, data = init20stem)
summary(i20stem_N)
Anova(i20stem_N)

as.data.frame(Anova(i20stem_N))
init20_stemN_thin <- emmeans(i20stem_N, pairwise ~ per_thin)

emmeans(i20stem_N, pairwise ~ per_thin*shd_trt, adjust="none")

i20stem_thin_ci = as.data.frame(init20_stemN_thin$emmeans)
```


# MATURITY 2020 (y+1)

```{r}

h20res <- read.csv("Resource Analysis_Har_2020.csv", header = T)

#inspect the first six rows, make sure it looks good. 
head(h20res) 

```

```{r}
har20res<- h20res %>%
  mutate(per_thin = as.factor(per_thin)) 
```


```{r}
har20stem <- har20res[har20res$tissue =="stem",]
 head(har20stem)
```

###  model 
```{r}
har20stem_N<- glm(per_n ~ per_thin * shd_trt * fert_trt, data = har20stem)

summary(har20stem_N)
Anova(har20stem_N)


as.data.frame(Anova(har20stem_N))
har20_stemN_thin <- emmeans(har20stem_N, pairwise ~ per_thin, adjust="none")
har20stem_thin_ci = as.data.frame(har20_stemN_thin$emmeans)

 emmeans(har20stem_N, pairwise ~ fert_trt)
 emmeans(har20stem_N, pairwise ~ shd_trt)

```

### plot 

```{r}

ggplot(har20stem_thin_ci, aes(per_thin, emmean, fill=per_thin))+
  geom_errorbar(data = har20stem_thin_ci, mapping = aes(ymin=lower.CL, ymax=upper.CL), width = 0.5, color="black")+
    geom_point(size=4, pch=21)+
  scale_x_discrete(name="Relative Fruit Load (y)", limits =c("66", "0"), labels = c("33%","100%"))+
  scale_y_continuous(name = "Stem % N at Fruit Maturity (y+1)", limits = c(1,2))+
  theme_bw()+
  scale_fill_manual(values = c( "#2ca25f", "#e5f5f9"))+
  theme(text = element_text(size = 16))+
  theme(legend.position = "none")

```








